Abstract Data Type Inference

### 2015-03-20 matching.tex
Matching schemes include read-only labelings.
Thorough examples for matching scheme inference.

TODO MOTIVATE THIS SECTION

@@

  We say an execution of read and write methods \emph{writes unique values} if
  the argument value to each write operation is unique. Consider the language
  whose executions write unique values, with the following predicates:
  \begin{align*}
    & \mathsf{write}(e,o) & \text{$o$ is a {\sf write} operation in $e$}, \\
    & \mathsf{read}(e,o) & \text{$o$ is a {\sf read} operation in $e$,} \\
    & \mathsf{eq}(e,o_1,o_2) & \text{$o_1$ and $o_2$ are the same operation, and} \\
    & \mathsf{veq}(e,o_1,o_2) & \text{$o_1$ and $o_2$ read/write the same values.}
  \end{align*}
  Given a valuation $\tup{x_\mathsf{w}, x_\mathsf{r}, y_\mathsf{w},
  y_\mathsf{r}, z_\mathsf{=}, z_\mathsf{v}}$ of the predicates above, we define
  the function $G(x_\mathsf{w}, x_\mathsf{r}, y_\mathsf{w}, y_\mathsf{r},
  z_\mathsf{=}, z_\mathsf{v})$ to be satisfied if and only if:
  \begin{align*}
    ( x_\mathsf{w} \land z_\mathsf{=} )
    \lor ( x_\mathsf{w} \land y_\mathsf{r} \land z_\mathsf{v} )
  \end{align*}
  Intuitively, this defines a simple matching scheme for which write
  operations match themselves, and read operations match the write which wrote
  the value read. In the case such a write exists, it is unique in any
  execution which writes unique values. The match is otherwise undefined.

@@

@@

  TODO SHOW THAT THE REGISTER LANGUAGE EXAMPLE WITHOUT ONE OF THE PREDICATES
  WILL LEAD TO UNSAT.

>>>>

TODO MOTIVATE THIS SECTION, AND ALSO ACCOUNT FOR THE READ-ONLY OPERATION
LABELING TOO.

@@

\begin{example}
  \label{ex:register-language}

  We say an execution of read and write methods \emph{writes unique values} if
  the argument value to each write operation is unique. Consider the language
  $L_\mathrm{reg}$
  whose executions write unique values, with the following predicates:
  \begin{align*}
    & \mathsf{w}(e,o) & \text{$o$ is a {\sf write} operation in $e$, and}, \\
    & \mathsf{v}(e,o_1,o_2) & \text{$o_1$ and $o_2$ read/write the same values.}
  \end{align*}
  We define the function $G(x_\mathsf{w}, y_\mathsf{w}, z_\mathsf{v})$ over
  valuations to the predicates above to be satisfied if and only if:
  \begin{align*}
    x_\mathsf{w} \land z_\mathsf{v}
  \end{align*}
  Intuitively, this defines a simple matching scheme for which write
  operations match themselves, and read operations match the write which wrote
  the value read. In the case such a write exists, it is unique in any
  execution which writes unique values. The match is otherwise undefined.

@@

  Consider again the language $L_\mathrm{reg}$ of read and write methods of
  Example~\ref{ex:register-language} with predicates ${\sf w}(e,o)$ and ${\sf
  v}(e,o_1,o_2)$ whose executions write unique values. The following table
  lists all possible predicate valuations $\tup{x_{\sf w}, y_{\sf w}, z_{\sf
  v}}$, and for each valuation a valid execution which excludes the positive
  valuation of ${\sf g}(x_{\sf w}, y_{\sf w}, z_{\sf v})$ in the satisfaction
  of $\textsc{Normalize}_F$, in the case such an execution exists.
  \begin{center}
    \begin{tabular}{ccc|ll}
      $x_\mathsf{w}$ & $y_\mathsf{w}$ & $z_\mathsf{v}$ & counterexample & reason \\
      \hline
      $0$ & $0$ & $0$ & ${\rm w}(1) \ {\rm r}(1) \ {\rm r}(1) \ {\rm w}(2) \ {\rm r}(2)$ & not unique \\
      $0$ & $0$ & $1$ & ${\rm w}(1) \ {\rm r}(1) \ {\rm r}(1)$ & not unique \\
      $0$ & $1$ & $0$ & ${\rm w}(1) \ \underline{{\rm r}(1) \ {\rm w}(2)} \ {\rm r}(2)$ & $\to_\mathrm{m}$ \\
      $0$ & $1$ & $1$ & $\underline{{\rm w}(1) \ {\rm r}(1)} \ {\rm r}(1)$ & $\to_\mathrm{m}$ \\
      $1$ & $0$ & $0$ & ${\rm w}(1) \ {\rm w}(2) \ {\rm w}(3) \ {\rm r}(3)$ & not unique \\
      $1$ & $0$ & $1$ & & \\
      $1$ & $1$ & $0$ & ${\rm w}(1) \ {\rm w}(2) \ {\rm w}(3)$ & not unique \\
      $1$ & $1$ & $1$ & & \\
    \end{tabular}
  \end{center}
  For example, the first valuation $000$ must be excluded since it allows the
  read ${\rm r}(2)$ to match both reads ${\rm r}(1)$, violating
  $\textsc{UniqueMatch}$. The second valuation $001$ must also be excluded,
  since it allows either read ${\rm r}(1)$ to match both itself and the other
  read ${\rm r}(1)$. The third valuation $010$ must be excluded since it allows
  the write ${\rm w}(2)$ to match the read ${\rm r}(1)$, in which case removing
  the match $\set{ {\rm r}(1), {\rm w}(2) }$ results in the invalid execution
  ${\rm w}(1)\ {\rm r}(2)$. Reasoning follows similarly for the other rows.
  In this way, any enumeration $F$ which includes the above executions
  will exclude all valuations except for $101$ and $111$, ultimately resulting
  in a satisfiable $\textsc{Normalize}_F$ in which ${\sf g}(x_{\sf w}, y_{\sf
  w}, z_{\sf v})$ is the same Boolean function $x_{\sf w} \land z_{\sf v}$
  given in Example~\ref{ex:register-language}.

@@

  Consider again the language $L_\mathrm{reg}$ of read and write methods of
  Example~\ref{ex:register-language}, yet this time without the predicate ${\sf
  v}(e,o_1,o_2)$. The following table lists all possible predicate valuations
  $\tup{x_{\sf w}, y_{\sf w}}$, and for each valuation a valid execution which
  excludes the positive valuation of ${\sf g}(x_{\sf w}, y_{\sf w})$ in the
  satisfaction of $\textsc{Normalize}_F$.
  \begin{center}
    \begin{tabular}{cc|ll}
      $x_{\sf w}$ & $y_{\sf w}$ & counterexample & reason \\
      \hline
      0 & 0 & ${\rm w}(1)\ {\rm r}(1)\ {\rm r}(1)$ & not unique \\
      0 & 1 & $\underline{{\rm w}(1)\ {\rm r}(1)}\ {\rm r}(1)$ & $\to_\mathrm{m}$ \\
      1 & 0 & ${\rm w}(1)\ {\rm w}(2)\ {\rm r}(2)$ & not unique \\
      1 & 1 & ${\rm w}(1)\ {\rm w}(2)$ & not unique \\
    \end{tabular}
  \end{center}
  Thus any enumeration $F$ including the executions above excludes every
  possible valuations of $\tup{x_{\sf w}, y_{\sf w}}$, resulting in an
  unsatisfiable $\textsc{Normalize}_F$.
